---
title: "Homework 1"
author: Rose Killian (rk3110)
date: 2021-09-25
output: github_document
---

First homework for p8105! 


First we load the needed libraries
```{r, message = FALSE}
library(tidyverse)
```


# Problem 1

Here's a **code chunk** that creates a _data frame_ with four different variables:

```{r}
df_1 = tibble(
  sample = rnorm(10),
  gr_than_zero = sample > 0,
  char = c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten"),
  factor = factor(c("yes", "no", "maybe", "yes", "no", "maybe", "yes", "no", "maybe", "yes"))
  )
```

Next we try to take the mean of each variable:

```{r}
mean(pull(df_1, sample))

mean(pull(df_1, gr_than_zero))

mean(pull(df_1, char))

mean(pull(df_1, factor))
```

It only works for the `sample` and `gr_than_zero` variables! This is because you can only take the mean of a numeric or logical variable. Let's apply the `as.numeric` function to some of our variables:

```{r, eval = FALSE}
as.numeric(pull(df_1, gr_than_zero))

as.numeric(pull(df_1, char))

as.numeric(pull(df_1, factor))
```

Treating our factor variable as numeric converts it to its respective 3 levels and we can now take its mean in addition to that of our numeric and logical variables. We still cannot take the mean of the character variable as treating it as numeric replaces all values with NA/missing values. By default, R cannot take the mean of a variable with missing values.


# Problem 2

Let's load in our data set:

```{r}
data("penguins", package = "palmerpenguins")
```

The variables in this data set are `r names(penguins)`. The mean body mass in grams is `r mean(pull(penguins, body_mass_g), na.rm = TRUE)` and the species of penguin included in the data set are `r levels(pull(penguins, species))`. The data set is `r nrow(penguins)` rows long and `r ncol(penguins)` columns wide. The mean flipper length of the data set is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)`mm.

Next let's make a plot of the data:

```{r, warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
geom_point() +
theme(legend.position = "bottom")

ggsave("scatter_plot.pdf")
```

